COVID Shutdown Regulations Negatively Impact Employment and Sales of Restaurants in Ontario

### 2021-02-25 outputs/paper/paper.Rmd
Commit for content structure

---
title: "COVID Shutdown Regulations Negatively Impact Employment and Sales of Restaurants in Ontario"
author: "Hong Shi, Hong Pan, Yixin Guan, Babak Mokri"
thanks: "Code and data are available at: https://github.com/honn-ishinn/restaurants_covid_shutdown"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Due to the world-wide COVID-19 pandemic, shutdown restrictions on businesses and facilities have limited economic activities into a recession. This paper examines the shutdown effect on restaurants by reopening a city in Ontario (Peterborough) and comparing its restaurant operating situation with that of another city (Brantford) after three months reopening. The survey result indicates that shutdown restrictions have negative impacts on restaurant employment, sales and future operation. However, the higher increased COVID cases in the reopened city sugguests the Ontario government to mindfully plan for futher policy decisions."
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
toc: FALSE

@@

&nbsp; &nbsp; To help recover from the current recession, we at Petit Poll have partnered with the Ontario government to examine the effects and scope of COVID shutdown on restaurants. We conducted an experiment by selecting two statistically similar cities in Ontario and investigated how reopening the dine-in services for restaurants in one city has prevailed over the state of continued shutdown in another city. We compared unemployment rates, restaurant profitability, and the expected financial and social impact in both circumstances. The results showed that general satisfaction, a sense of purpose, hope, and profitability increase among the businesses that were allowed to re-open. The findings can be generalized to other similar businesses and their associated families and business partners in that city as well. In addition, the governmental support programs for small businesses in both cities contributed substantially to the social wellbeing and recovery pathway of all businesses. However, the higher comparative rate of COVID cases in the city under treatment resulted in general anxiety among residents, triggered the social health advocates and activists to criticize the government reopening effort, and undermined the projected benefits. We then came across the trade-off between opening or closing the public parts of restaurants - or other similar businesses, for which we suggested utilizing SSM (Soft System Methodology) to collect, consider and accommodate all residents' perspectives and potential viewpoints in society.

@@

&nbsp; &nbsp; However, due to current COVID pandemic, any loosening of shutdown restrictions increases chances of COVID exposures, resulting in challenges on public health and healthcare of reopened regions. It is highly risky and hasty to reopen a large area of Ontario regions for the sole purpose of our experiment. Therefore, we mindfully chose Peterborough, a relative small size Single-tier^[Ontario's Municipal Act, 2001 defines a single-tier municipality as "a municipality, other than an upper-tier municipality, that does not form part of an upper-tier municipality for municipal purposes"] city 125 kilometers northeast of Toronto, as the only region to reopen in Ontario and then compared its restaurant operating situations with those in Brantford, another Single-tier city 105 kilometers southwest of Toronto with similar population. Since demographics, land areas and more importantly, restaurant operating situations between these two cities are quite similar with each other^[Approach to examine the similarities of these two cities are further introduced in the discussion section], we consider these two cities as valid comparison cities that could yield an accurate estimate of restaurant businesses shutdowns. In our experiment, we assigned restaurants in Peterborough as the treatment group and those in Brantford as the control group. We also ensured restaurants within these two cities to have equal opportunities receiving government support (e.g. subsidies, grants, etc.) to estimate the true impact of shutdown effects on restaurant businesses after three-month reopening in Peterborough.

@@

## Sampling Methods

@@

## Survey Results and Policy Implication
-	Staffing implications 
•	Policy implications: The Government of Ontario should continue the existing supports provided to restaurants during the lockdown and consider additional support measures to mitigate the risk of increased unemployment. Reopening the restaurants may lead to increased employment in the corresponding area. 
-	Restaurant sales implementation 
•	The Government should expect better sales for reopened restaurants. However, it is important to acknowledge that economy would take a long time to reach pre-pandemic performance. Continued support to restaurants may still be needed. 
-	Future projects of restaurants  
•	Rebuilding business confidence is a crucial aspect of short-term business recovery and long-term economic stability. The Government should carefully monitor business projection and equip businesses with tailored supports based on their recovery needs. 
-	Striking a balance between COVID-19 cases
•	While reopening leads to increased employment, improved restaurant sales, positive business confidence, and ultimately increased government revenue, the impact of the reopening of restaurants needs to be assessed against the situation of the pandemic. By closely monitoring the development of confirmed COVID-19 cases during the intervention period, we found a more rapid spread of COVID-19 cases in Peterborough compared to Brantford. Although more quantitative and qualitative research needs to be conducted to determine if the surge is related to the reopening intervention, the results serve as an alarming-clock for policymakers and the public. 
•	The government should strike a balance between economic recovery and public health and safety. The health and safety of Ontarians should always be the top priority for the government. The Government should carefully monitor and assess the impact of COVID-19 cases and adjust its reopening policies accordingly. 
•	In the case of reopening and possible reclosing of restaurants in a region, the government should clearly and effectively communicate the risks and harms of COVID-19 resurgence and its willingness and devotion to addressing business concerns.

>>>>

---
title: "COVID Shutdown Regulations Negatively Impact Employment and Sales of Restaurants in Ontario"
author: "Hong Shi, Hong Pan, Yixin Guan, Babak Mokri"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Due to the world-wide COVID-19 pandemic, shutdown restrictions on businesses and facilities have limited economic activities into a recession. This paper examines the shutdown effect on restaurants by reopening a city in Ontario (Peterborough) and comparing its restaurant operating situation with that of another city (Brantford) after three months reopening. The survey result indicates that shutdown restrictions have negative impacts on restaurant employment, sales and operating confidence. However, the higher increased COVID cases in the reopened city sugguests the Ontario government to mindfully plan for futher policy decisions."
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
toc: FALSE

@@

&nbsp; &nbsp; To help recover from the current recession, we at Petit Poll have partnered with the Ontario government to examine the effects and scope of COVID shutdown on restaurants. We conducted an experiment by selecting two single tier cities, namely Peterborough and Brantford, ^[Ontario's Municipal Act, 2001 defines a single-tier municipality as "a municipality, other than an upper-tier municipality, that does not form part of an upper-tier municipality for municipal purposes"] with similar restaurant operating situations in Ontario and investigated how reopening the dine-in services for restaurants in Peterborough has prevailed over the state of continued shutdown in Brantford through surveys. The results showed that continuing shutdown has negatively impact the employment, sales volume and operating confidence on restaurant businesses in Brantford. In addition, the governmental support programs contributed substantially to restaurant businesses in both reopened and shutdown cities. However, the higher increased COVID cases in Peterborough demonstrates the drawback of the reopening, and the Ontario government should hold a clear measure between public health and economic recovery before further policy implementations.

&nbsp; &nbsp; We organize he rest of paper as follows: In Data section, we first discuss interventions of the experiment. Next, we introduce sampling methods and survey methodologies to collect the data. And then, we discuss main findings from survey results. In Discussion section, we would elaborate on concerns about the intervention on city selections, sampling methods and survey methodologies. And we also propose our thoughts on further policy implications. In Appendix, we include screenshots of survey contents.^[ Code and data are available at the GitHub repo: https://github.com/honn-ishinn/restaurants_covid_shutdown"]

@@

&nbsp; &nbsp; However, due to current COVID pandemic, any loosening of shutdown restrictions increases chances of COVID exposures, resulting in challenges on public health and healthcare of reopened regions. It is highly risky and hasty to reopen a large area of Ontario regions for the sole purpose of our experiment. Therefore, we mindfully chose Peterborough, a relative small size Single-tier city 125 kilometers northeast of Toronto, as the only region to reopen in Ontario and then compared its restaurant operating situations with those in Brantford, another Single-tier city 105 kilometers southwest of Toronto with similar population. Since demographics, land areas and more importantly, restaurant operating situations between these two cities are quite similar with each other^[Approach to examine the similarities of these two cities are further introduced in the discussion section], we consider these two cities as valid comparison cities that could yield an accurate estimate of restaurant businesses shutdowns. In our experiment, we assigned restaurants in Peterborough as the treatment group and those in Brantford as the control group. We also ensured restaurants within these two cities to have equal opportunities receiving government support (e.g. subsidies, grants, etc.) to estimate the true impact of shutdown effects on restaurant businesses after three-month reopening in Peterborough.

@@

## Sampling Method and Survey Methodology

@@

## Survey Results and Policy Implication
-	Staffing implications 
*	Policy implications: The Government of Ontario should continue the existing supports provided to restaurants during the lockdown and consider additional support measures to mitigate the risk of increased unemployment. Reopening the restaurants may lead to increased employment in the corresponding area. 
-	Restaurant sales implementation 
*	The Government should expect better sales for reopened restaurants. However, it is important to acknowledge that economy would take a long time to reach pre-pandemic performance. Continued support to restaurants may still be needed. 
-	Future projects of restaurants  
*	Rebuilding business confidence is a crucial aspect of short-term business recovery and long-term economic stability. The Government should carefully monitor business projection and equip businesses with tailored supports based on their recovery needs. 
-	Striking a balance between COVID-19 cases

```{r covidcase, fig.align='center', fig.cap="Total Confirmed COVID-19 cases in Peterborough and Brantford", fig.width=8, fig.height=4, echo= FALSE, message=FALSE, warning=FALSE}
  ggplot(data=covid_tracking,aes(x=track_date))+
    geom_smooth(aes(y=covid_cases_p,color="Peterborough"))+
    geom_smooth(aes(y=covid_cases_b,color="Brantford"))+
    theme_minimal()+
    scale_color_manual("City",breaks = c("Peterborough","Brantford"),
                       values = c("red","blue"))+
    xlab("Tracking date")+
    scale_y_continuous("Cumulative number of confirmed cases",limits=c(0,800))+
    labs(title = " Confirmed COVID-19 cases in Peterborough and Brantford")+
    scale_x_date(date_labels = "%Y %b %d")+
   theme(plot.title = element_text(hjust = 0.5)) +
    guides(color=guide_legend(override.aes=list(fill=NA)))
```

*	While reopening leads to increased employment, improved restaurant sales, positive business confidence, and ultimately increased government revenue, the impact of the reopening of restaurants needs to be assessed against the situation of the pandemic. By closely monitoring the development of confirmed COVID-19 cases during the intervention period, we found a more rapid spread of COVID-19 cases in Peterborough compared to Brantford (Figure). Although more quantitative and qualitative research needs to be conducted to determine if the surge is related to the reopening intervention, the results serve as an alarming-clock for policymakers and the public. 
*	The government should strike a balance between economic recovery and public health and safety. The health and safety of Ontarians should always be the top priority for the government. The Government should carefully monitor and assess the impact of COVID-19 cases and adjust its reopening policies accordingly. 
*	In the case of reopening and possible reclosing of restaurants in a region, the government should clearly and effectively communicate the risks and harms of COVID-19 resurgence and its willingness and devotion to addressing business concerns.
